---
title: "HackathonP3Problem"
author: "Gazal"
date: "22 September 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r global `functions}
library(graphics)
library(ggplot2)

getmode <- function(v)
{
  uniqv <- unique(v)
  uniqv[which.max(tabulate(match(v, uniqv)))]
}

color_v=c("gray37","burlywood3","peachpuff4",
          "chocolate1","darkgoldenrod1","coral2",
          "mediumorchid1","cadetblue")

```

# Movie Database

## About Database
```{r dataset}
  library(pastecs)

  movie_df = read.csv("D:/STUDY/hackathon/hacaktontopic/Data\ Set\ -\ movies_withNullValues.csv")
  print(class(movie_df))
  print(summary(movie_df))
  print(str(movie_df))
  print(colnames(movie_df))

  print(stat.desc(movie_df))
```

##Converting False Factors into integer manually

```{r replace false factors into integer}

# domgross
# intgross
# budget_2013.
# domgross_2013.
# intgross_2013.

movie_df$domgross = as.numeric(paste(movie_df$domgross))
movie_df$intgross = as.numeric(paste(movie_df$intgross))
movie_df$budget_2013. = as.numeric(paste(movie_df$budget_2013.))
movie_df$domgross_2013. = as.numeric(paste(movie_df$domgross_2013.))
movie_df$intgross_2013. = as.numeric(paste(movie_df$intgross_2013.))

#print(str(movie_df))
print(summary(movie_df))
```
## understanding missing values

```{r explorating null values in dataset, echo=FALSE}
library(mice)
library(VIM)

data_pattern = md.pattern(movie_df)

print(data_pattern)

imputation_plot = aggr(movie_df,
                       col = color_v,
                       numbers = TRUE,
                       sortVars = FALSE,
                       labels = names(movie_df),
                       cex.axis = 0.9,
                       gap = 2,
                       ylabs = c("Missing Data","Pattern"))
print(imputation_plot)
```

```{r missing values percentage}
  pMiss <- function(x){sum(is.na(x))/length(x)*100} 
  apply(movie_df,2,pMiss) 
```

##1. Summative Analysis describing the number of
  a. Columns
  b. rows
  c. names of different columns
  d. The class types of the columns

```{r q1}

# a: columns

  total_cols = ncol(movie_df)
  cat("\n Number of columns: ", total_cols)
  
# b: rows
  
  total_rows = nrow(movie_df)
  cat("\n Number of rows: ", total_rows)
  
# c: names of different columns
  
  total_col_names = colnames(movie_df)
  cat("\n Number of column names: \n")
  print(total_col_names)
  
# The class types of the columns
  
  print("\n Class types: ")
  total_class_types = vector()
  for(i in 1:total_cols)
  {
    total_class_types[i] = class(movie_df[1,i])
    cat("\n Type of ", total_col_names[i]," is ", total_class_types[i])  
  }

```

##2. Check the Data Quality of the data. If you realize there are null values in it, use Data Imputation for filling the data. Justify the methodolog
```{r association, message=FALSE, warning=FALSE}

# assoication between code -> year and binary
  code_year = as.numeric(substr(movie_df$code, 1, 4))
  code_binary = as.factor(substr(movie_df$code, 5, 8))
  
  plot(code_year,movie_df$year, type= "l", col = color_v)
  plot(as.numeric(code_binary),
       as.numeric(movie_df$binary), type = "l", col = color_v)

# assoication between year -> period-code
  plot(movie_df$year, movie_df$period.code, type = 'b', col = color_v)
  
# assoication between year -> decade-code
  plot(movie_df$year, movie_df$decade.code, type = 'b', col = color_v)
  
```

```{r q2, message=FALSE, warning=FALSE}
#imputing period code

dat = movie_df
dat_imp<-dat
for (i in 1:nrow(dat)) {
   if (is.na(dat[i, 14])) {
      if(dat$year[i]>=1985 & dat$year[i]<1990){
       dat_imp[i, 14] =6
      }else if(1980<=dat$year[i] & dat$year[i]<1985){
       dat_imp[i, 14] =7}else if(1975<=dat$year[i] &dat$year[i]<1980){
         dat_imp[i, 14] =8}else if(1970<=dat$year[i] & dat$year[i]<1975){
         dat_imp[i, 14] =9}else{dat_imp[i, 14] =1}
     
   }
}


#imputing decade code
for (i in 1:nrow(dat)) {
   if (is.na(dat[i, 15])) {
      if(dat$year[i]>=2010){
       dat_imp[i, 15] =1
      }else if(1980<=dat$year[i] & dat$year[i]<1990){
       dat_imp[i, 15] =4}else if(1970<=dat$year[i] &dat$year[i]<1980){
         dat_imp[i, 15] =5}
     
   }
}
#convert intgross to numeric
dat_imp$intgross<-as.numeric(paste(dat$intgross))

#median intgross
intgross_median<-median(dat_imp$intgross,na.rm = TRUE)

#convert intgross_2013 to numeric
dat_imp$intgross_2013.<-as.numeric(paste(dat$intgross_2013.))

#mmedian intgross_2013
intgross_2013_median<-median(dat_imp$intgross_2013.,na.rm = TRUE)

#convert domgross to numeric
dat_imp$domgross<-as.numeric(paste(dat$domgross))

#median domgross
domgross_median<-median(dat_imp$domgross,na.rm = TRUE)

#convert domgross_2013 to numeric
dat_imp$domgross_2013.<-as.numeric(paste(dat$domgross_2013.))

#median domgross
domgross_2013_median<-median(dat_imp$domgross_2013.,na.rm = TRUE)

#CODE COVERSION
dat_imp$code<-as.character(dat_imp$code)

#median of budget
budget_median<-median(dat_imp$budget,na.rm = TRUE)
#imputing remaining data
for (i in 1:nrow(dat_imp)) {
   if (is.na(dat_imp[i, 9])) {
     dat_imp[i, 9] =intgross_median 
   }
  if (is.na(dat_imp[i, 8])) {
     dat_imp[i, 8] =domgross_median 
     
  }
   if (is.na(dat_imp[i,12])) {
     dat_imp[i,12 ] =domgross_2013_median 
     
   }
   if (is.na(dat_imp[i, 13])) {
     dat_imp[i,13 ] =intgross_2013_median
     
   }
  if (is.na(dat_imp[i,10 ])) {
     dat_imp[i,10]<-paste(dat_imp$year[i],dat_imp$binary[i],sep= "")
   
  }
   if (is.na(dat_imp[i, 7])) {
     dat_imp[i,7 ] =budget_median
     
   }
  
}
dat_imp$code <- as.factor(dat_imp$code)
print(summary(dat_imp))

movie_df = dat_imp
print(stat.desc(movie_df))
  
```

## after replacing null values (data imputation)

```{r explorating null values in dataset after, echo=FALSE}
data_pattern = md.pattern(movie_df)

print(data_pattern)

imputation_plot = aggr(movie_df,
                       col = color_v,
                       numbers = TRUE,
                       sortVars = FALSE,
                       labels = names(movie_df),
                       cex.axis = 0.9,
                       gap = 2,
                       ylabs = c("Missing Data","Pattern"))
print(imputation_plot)
```
## 3. Do Univariate Analysis of all the variables present

  Type of Variables:
  1) Quanititative -> Numeric
  2) Qualitative -> Categorical(Factor)
  
### Understanding numeric data

```{r continuous, message=FALSE, warning=FALSE}

# dataset of continuous variables of data
int_df <- Filter(is.numeric, movie_df)
#print(head(int_df))

# removing SL_NO and Mobile no as they are not very much useful for analysis
int_df <- int_df[c(-1,-2)]

# continuous variable colnames
sdf_cols_int <- colnames(int_df)

print(sdf_cols_int)

for(i in 1:ncol(int_df))
{
  # remove null from column
  column_name = sdf_cols_int[i]
  cat("\n\n column : ",column_name)
  
  column_int_df = int_df[i]
  column_int_df = column_int_df[!apply(is.na(column_int_df) |
                                             column_int_df == "", 1, all),]
  
  # count of total values
  total_count_column = nrow(int_df[i])
  
  # count of null values  
  column_int_df_null_count = total_count_column - length(column_int_df)

  # percentage of null in column
  column_null_perc = column_int_df_null_count/total_count_column
  cat("\n null value percentage : ", column_null_perc)
  
  # range of column
  column_range = range(column_int_df)
  cat("\n range : ", column_range)
  
  # printing quantile of column
  column_quantile = quantile(column_int_df)
  cat("\n quantile : ", column_quantile)
  
  # printing minimum of column
  column_min = min(column_int_df)
  cat("\n minimum : ", column_min)
  
  # printing maximum of column
  column_max = max(column_int_df)
  cat("\n maximum : ", column_max)
  
  # printing mean of column
  column_mean = mean(column_int_df)
  cat("\n mean : ", column_mean)
  
  # printing median of column
  column_median = median(column_int_df)
  cat("\n median : ", column_median)
  
  # printing mode of column using user defined function
  column_mode = getmode(column_int_df)
  cat("\n mode : ", column_mode)
    
  # Median Absolute Deviation
  column_mad = mad(column_int_df)
  cat("\n median absolute deviation : ", column_mad)
  
  # variance
  column_variance = var(column_int_df)
  cat("\n variance : ", column_variance)
  
  # standanrd deviation
  column_sd = sd(column_int_df)
  cat("\n standard deviance : ", column_sd)
  
  # understanding scattered data
  plot(column_int_df, xlab = column_name, ylab = "Values",
             col = color_v,
             main = paste("Scatter Plot for", column_name))
  # understanding frequency distribution of values
  hist(column_int_df, xlab = column_name, ylab = "Frequency",
             col = color_v,
             main = paste("Histogram for", column_name))
  
  # undertstanding outliers using boxplot
  boxplot(column_int_df, xlab = column_name,
               col=color_v)
  

}

```

### Understanding numeric data

```{r categorical, message=FALSE, warning=FALSE}
library(plyr)

factor_df <- Filter(is.factor, movie_df)

# removing name and email id as they are not very much useful for analysis
factor_df <- factor_df[c(-1,-2)]

# factor variable colnames
sdf_cols_factor <- colnames(factor_df)

print(sdf_cols_factor)

for(i in 1:length(sdf_cols_factor))
{
  column_name = sdf_cols_factor[i]
  cat("\n\n Factor : ")
  print(column_name)
  column_factor_df = as.factor(factor_df[,i])

  # count of total values
  total_count_column = nrow(factor_df[i])
  
  cat("\n levels : ")
  print(levels(column_factor_df))
  
  cat("\n number of levels : ")
  print(nlevels(column_factor_df))
  
  cat("\n Orders? : ")
  print(is.ordered(column_factor_df))
  
  mode_fac = getmode(column_factor_df)
  cat("\n Mode : ")
  print(mode_fac)
  
  count_fac = count(column_factor_df)
  print(count_fac)
  
  # understanding frequency
  barplot(prop.table(table(column_factor_df)))  
  
  # understanding relative frequency of values
  
  labels <- count_fac$x
  x <- count_fac$freq
  
  pie(x, labels, main = column_name, col = rainbow(length(x)))

}

```

## 4. The target variable in this dataset is the attribute named "binary" which has two predefined variables FAIL and PASS. Do a bivariate analysis of this variable with all the other attributes. Also give your inferences from the graph generated.

```{r bivariate plots, echo=FALSE, message=FALSE, warning=FALSE}
for(i in 1:total_cols)
{
    plot( movie_df$binary, movie_df[,i],
       xlab = "Binary", ylab = total_col_names[i], col = color_v)
  
       legend(2000,9.5, legend = levels(movie_df$binary))
  
     plot(movie_df[,i], movie_df$binary,
       ylab = "Binary", xlab = total_col_names[i], col = color_v)
        
       legend(2000,9.5, legend = levels(movie_df$binary))
}


```

## 5. Draw a multivariate graph for all the variables

```{r multivariate graph, echo=FALSE, message=FALSE, warning=FALSE}
  plot(movie_df)
```

## 6. Also draw a graph between year and budget

```{r q6, message=FALSE, warning=FALSE}


plot(movie_df$year,movie_df$budget)
plot(movie_df$year,movie_df$budget,
           type = 'l')

#ggplot(data = movie_df, aes(x = year, y = budget)) + geom_hline()

# ggplot( movie_df, aes(movie_df$year) ) + 
#   geom_line( aes( y = movie_df$budget ) ) +
#   labs( title = "graph")


ggplot( movie_df, aes( x= movie_df$year, y = movie_df$budget ) ) + geom_smooth(model = "lm")

ggplot( movie_df, aes( x= movie_df$year, y = movie_df$budget) ) + geom_line()




```

## 7. Apply PCA on this dataset after removing categorical variables

Prerequisite for applying PCA:
1) Only Numeric Data present
2) no Null values are present

```{r pca, message=FALSE, warning=FALSE}

movie_NA_df = data.frame()
movie_df <- Filter(is.numeric, movie_df)
movie_NA_df <- vapply(as.data.frame(movie_df), function(x)
                    replace(x, is.na(x), median(x,na.rm=TRUE)),
                    FUN.VALUE=numeric(nrow(movie_df)) )

print(head(movie_NA_df))

#attitude_PCA = princomp(attitude_df)
movie_NA_PCA = princomp(movie_NA_df)

#print(names(attitude_PCA))
print(names(movie_NA_PCA))

#biplot(attitude_PCA)

biplot(movie_NA_PCA, col = color_v)

sd_pca<-movie_NA_PCA$sdev

# get variance of pca
 var_pca = sd_pca ^ 2
 
 # variance proportion respective to sum
 proportion_var_pca = var_pca/sum(var_pca)
 
 # variance plot
 plot(proportion_var_pca,
      xlab = "Principal Components",
      ylab = "Proportion of Variance Explained",
      type = "b",
      col = color_v)
 
 # cumulative variance plot
 plot(cumsum(proportion_var_pca),
      xlab = "Principal Components",
      ylab = "Proportion of Variance Explained",
      type = "b",
      col = color_v)

 # score plot
 plot(movie_NA_PCA$scores[,1], movie_NA_PCA$scores[,2], col = color_v)
```


